Wireless service accounting based on Ethereum state channels - Master Thesis

### 2019-10-26 Paper.tex
Add static analysis tools outcome

\begin{abstract}
  \begin{center}
    This document contains authors' ``\textit{Blockchain
      technologies master}''\footnote{UPC School
      of Professional \& Executive Development}
    final thesis. The project described in the document
    consists on a technological proof
    of concept of an internet access service consumption
    using a state channel smart contract as
    mean of payment. After a brief
    introduction, there's a first motivation of
    the use case included;
    in the following sections
    there's a detailed explanation of
    the architecture, implementation and
    security considerations;
    after that it is described a more
    general approach about using blockchain technologies
    and state channels to provide
    mechanisms to reduce advanced wireless
    (i. e. 5G) services market friction between neutral host
    providers, virtual operators and operators
    purchasing licenses of spectrum bands to nations' governments;
    ideally achieving an unprecedented transparency guarantee to consumers and
    a fully GDPR compliant
    schema relying on a self-sovereign identity system.
  \end{center}
\end{abstract}

@@

\begin{center}
\Huge{\textbf{DISCLAIMER}}
\vspace{1cm}

\Large
All software described and linked in this document is open source
under the MIT license; so anyone can freely use it and make
modifications to the source code to fulfill their use
cases. But if you do so, please take into account that the software
at the time of this writing is merely
a proof of concept and is not properly audited; although
authors tried to follow best practices regarding smart contract,
front-end and back-end design and implementation there are still
some trade-offs and recognizable security issues
described in this document's
security considerations section. Please take this into account
when using and/or modifying this software and be conscious
that authors do not provide any guarantee in case of
security issues, economical loses or regulation infringement.
\end{center}

@@

\begin{verbatim}
MIT License

@@

@@

https://github.com/ethereum-internet-access/mythril
Firstly, we analyzed the local source code:

@@

Secondly, we analyzed on-chain smart contract.
Because the smart contract was deployed on Ropsten test-net we executed:

@@

@@

To take into account the problem illustrated in previous paragraph
we can take as a basis a model of self-sovereign identity\cite{allen}
and in particular the one recently announced by the authorities
of Catalonia's government\cite{identicat} that will provide credentials
to citizens using DLT technologies as a Public Key Infrastructure; this
model could allow in some near future to give the citizen the option
of registering identity in such a way that her/his credentials can be linked
to some certified database that could be queried only by a judge
requirement and guarantee the users identities are recognized
%% and at the
%% same time hidden from service providers, discharging them from
%% any GDPR responsibilities apart from the intrinsic service provision
%% metadata.

@@

World we live today is interconnected and wireless
internet access has a fundamental role. In this project
a Proof of Concept has been build. This PoC based in
commodity free hardware asset (Raspberry Pi) can serve
as an example of end-user wireless provision pay-as-you-go
using a state channel Solidity smart contract; KyC issues
and identity consideration have been discussed also leading
to the conclusion that complementing this approach with
what a self-sovereign identity can provide there's no need
to a telecommunication operator to perform a KyC process
in order to provision a service to a client.

>>>>

\begin{abstract}
%  \begin{center}
    This document contains authors' ``\textit{Blockchain
      technologies master}''\footnote{UPC School
      of Professional \& Executive Development}
    final thesis. The project described in the document
    consists on a technological proof
    of concept of an internet access service consumption
    using a state channel smart contract as
    mean of payment. After a brief
    introduction, there's a first motivation of
    the use case included;
    in the following sections
    there's a detailed explanation of
    the architecture, implementation and
    security considerations;
    after that it is described a more
    general approach about using blockchain technologies
    and state channels to provide
    mechanisms to reduce advanced wireless
    (i. e. 5G) services market friction between neutral host
    providers, virtual operators and operators
    purchasing licenses of spectrum bands to nations' governments;
    ideally achieving an unprecedented transparency guarantee to consumers and
    a fully GDPR compliant
    schema relying on a self-sovereign identity system.

%  \end{center}
\end{abstract}

@@

% \begin{center}
\Huge{\textbf{DISCLAIMER}}
\vspace{1cm}

\Large
All software described and linked in this document is open source
under the MIT license; so anyone can freely use it and make
modifications to the source code to fulfill their use
cases. But if you do so, please take into account that the software
at the time of this writing is merely
a proof of concept and is not properly audited; although
authors tried to follow best practices regarding smart contract,
front-end and back-end design and implementation there are still
some trade-offs and recognizable security issues
described in this document's
security considerations section. Please take this into account
when using and/or modifying this software and be conscious
that authors do not provide any guarantee in case of
security issues, economical loses or regulation infringement.
% \end{center}

@@

\vspace*{5cm}
\normalsize
\begin{verbatim}
MIT License

@@

\subsection{Static analysis}

@@

https://github.com/ethereum-internet-access/mythril
Firstly, the local source code was analyzed:

@@

Secondly, on-chain smart contract was analyzed.
Because the smart contract was deployed on Ropsten test-net,  executed:

@@

\subsubsection{Analysis tools outcome}

\paragraph{Dependence on predictable environment variable}

Both Mythrill and Manticore have reported a low severity warning located on function
\textit{claimTimeOut}, due to the use of the special variable \textit{block.timestamp}, -refered
to as \textit{now}.

Mythril states:

\textit{A control flow decision is made based
on a predictable variable. The block.timestamp environment variable is used in to determine
a control flow decision. Note that the values of variables like coinbase, gaslimit,
block number and timestamp are predictable and can be manipulated by a malicious miner.
Also keep in mind that attackers know hashes of earlier blocks. Don't use any of those
environment variables for random number generation or to make critical control flow decisions.}

This warning is considered assumable due to this two facts: 1) the tiny incentive for the
malicious miner, who merely could steal the few \textit{weis} value of a channel in favour
of the \textit{user}, and 2) the uncertainty of reaching the goal of claiming the refund
before the contract owner closes it.

\paragraph{Multiple Calls in a Single Transaction}

Mythrill also reports this low severity warning located on function \textit{closeChannel}:

\textit{Multiple calls are executed in the same
transaction. This call is executed after a previous call in the same transaction.
Try to isolate each call, transfer or send into its own transaction.}

A workaround in order to avoid this issue couldn't be found, due to the fact that both
transfers must be done and for the low amount of value
allowed in implied transactions authors preferred
the push over pull pattern to avoid on-chain transaction overhead.
In the \textit{Writing a Simple Payment Channel} example available in
\href{https://solidity.readthedocs.io/en/latest/solidity-by-example.html#micropayment-channel}{Micropayment Channel},
a \textit{selfdestruct} is proposed. But in this project, a single smart contract is expected
to support several channels, rather than deploying a fresh contract in every connection.
Furthermore, both calls to \textit{transfer} happen after any changes to state variables,
so the contract is not vulnerable to a reentrancy exploit.

\paragraph{Beyond the scope of a PoC}

Some improvements should be considered prior to taking the \textit{StateChannel.sol}
implementation beyond the scope of this PoC:
\begin{itemize}
\item a setter function for \textit{pricePerSecond} should be in place. However, for the sake of
fairness, nobody should be allowed to update the price after the channel is opened. So therefore,
a new atribute storing the current channel price should be added to the \textit{ChannelData} struct.
\item minimal price allowed in channel openning should be a variable, rather than a hardcoded value.
\item only accepting ETH is a hard constrain that probably will frustrate any chance of a massive
adoption of this offering: ERC20 tokens should be considered.
\end{itemize}

@@

To take into account the problem illustrated in previous paragraph
a model of self-sovereign identity\cite{allen} can be taken as a basis
and in particular the one recently announced by the authorities
of Catalonia's government\cite{identicat} that will provide credentials
to citizens using DLT technologies as a Public Key Infrastructure; this
model could allow in some near future to give the citizen the option
of registering identity in such a way that her/his credentials can be linked
to some certified database that could be queried only by a judge
requirement and guarantee the users identities are recognized
%% and at the
%% same time hidden from service providers, discharging them from
%% any GDPR responsibilities apart from the intrinsic service provision
%% metadata.

@@

World today is interconnected and wireless
internet access has a fundamental role. In this project
a Proof of Concept has been build. This PoC based in
commodity free hardware asset (Raspberry Pi) can serve
as an example of end-user wireless provision pay-as-you-go
using a state channel Solidity smart contract; KyC issues
and identity consideration have been discussed also leading
to the conclusion that complementing this approach with
what a self-sovereign identity can provide there's no need
to a telecommunication operator to perform a KyC process
in order to provision a service to a client.
